Using tours to visually investigate properties of new projection pursuit indexes with application to problems in physics

### 2019-06-05 review/response.Rmd

## Author's response to reviews

The authors wish to thank the reviewers for the careful reading and comments on our first version. This is the detailed information on how we have addressed each comment from the reviewers.

### Associate Editor:

@@

The submission has been evaluated by two reviewers. The manuscript is well written and organized. I have some more philosophical issues in addition to those raised by the two reviewers.

The authors need to convince readers (users) this kind of environment is good for daily EDA for the real world of big data/data science.

**We have revised the introduction to improve the motivation to make this clear.** XXX CHECK THIS

@@

**We have analysed an example that is high-dimensional but think that it would be too distracting for the reader of this paper. It is an analysis of physics modeled fits, using initial ideas developed in this paper. There the dimensionality was around 60, and it worked best to reduce the dimensionality using PCA as a pre-processing step, because this explained most of the variability. The tour was applied on the reduced dimensionality, and the work is published in EPJC (cite).**

Who are the target users of this environment? Do you recommend naive users (users without much EDA experience) to get hand on the proposed system?

How do you encourage users to try your system?

**The target audience consists of two groups. First we address people interested in developing new index functions, who will find the introduced methods useful to diagnose new indexes as presented in the paper. The methods to do so are now easily accessible via the spinebil package. The second group would be people interested in applying the considered indexes to their datasets. This can be done directly using the tourr package, but we also encourage users unfamiliar with the system to try this through the GUI available in the galahr package. Both packages are now referenced in the discussion (or should it be mentioned somewhere else?).**

@@

**The methods are implemented in an R package, which makes them readily available to all. The code for this paper is all available on github, and we have now made this repo public. It depends on three other packages, mbgraphics, binostics, minerva, which are now updated and on CRAN. We have coded the diagnostic techniques into a new package, and will submit this to CRAN.**

### Reviewer #1:

Review report for "Using tours to visually investigate properties of new projection pursuit indexes with application to problems in physics"

@@

1) Authors conducted the simulation study to evaluate the selected PPIs based on the five criterions and the results were summarized into Table 3. On the other hand, authors may consider another table that summarizes the advantages/disadvantages/characteristics of these mathematical indexes from the previous literatures. This review is helpful for users to comprehend these indexes' properties.

**This comparison was the subject of the thesis referenced in the introduction (Grimm (2016)). We have now added a brief summary of these findings at the end of section 2.2.**

2) According to the aims & scope of the journal, Computational Statistics, this article should keep the content and discussion relevant more focused on the computational aspects.

**We are really not sure what is being requested here. We have tried to focus the content in the paper on the compututational aspects of the work, and have kept it concise and to the point. The paper is at a reasonable length. The appendix has some details that broaden the discussion. If there are more parts that the reviewer suggests should be shifted to an appendix, we would be happy to do that.**

@@

3) Although the target of this article focuses on the dataset with two-dimensional structure (d=2), it is worth to discuss the possibility (or give some suggestions) of applying the proposed procedure to the situation where d > 2.

**We thank the referee for the suggestion. There are many index functions available for d>2, and the ones in the tourr package work for arbitrary dimension. The work in this paper focuses on studying a selection of indexes designed for 2D. However, the diagnostics cold be applied generally to any dimension projection index. We have commented on this in the discussion.** XXX CHECK THIS.

@@

**Mention holes/cmass with reference in Sec. 2.2, but keep rescaling information in the appendix?**

8) Check the spelling of words for clarity and consistency, e.g, splines2D/splines2d; convex/1-convex, MIC/MICe, and so on.

**We have made the spelling consistent where referring to the same index, and provided explanations where the different spelling indicates differences in index calculation.**

### Reviewer #2:

Summary of the contribution:
The authors present a simulation study to evaluate various projection pursuit indexes for detecting specific bivariate patterns in the data. Moreover, two data sets from physics are examined using PPI.

### 2019-06-07 scagnostics.Rmd
Added summary of Katrins thesis.

\textcolor{red}{Here add short summary of comparisons between these functions for variable selection, taking results from Katrins thesis.}

>>>>

\textcolor{red}{A comparison between these indexes for the purpose of variable selection was discussed in @Grimm2016, and can be summarised as follows. The scagnostics measures are flexible and calculating the full set of measures provides useful guidance in variable selection, however they are found to be highly sensitive to outlying points and sample size (as a consequence of the binning). Both splines2D and dcor2D are found to be robust in this respect, but splines2D is limited to functional dependence, while dcor2D is found to take large values only in scenarios with large linear correlation. Finally the mutual information based index functions are found to be flexible, but are sensitive to the sample size and often take relatively large values also when no association is present. A comparison of MIC and dcor2D was also presented in @2014arXiv1401.7645S.}
